### 2018-10-01 BS_figures.tex
Add SI figures for raw data

% \documentclass{article}

% \usepackage{graphicx}
% \usepackage{float}

% \begin{document}
\maxdeadcycles=200
\begin{figure}[htbp]
  \centering
  {\bfseries \sffamily 1T-TiO$_{2}$}\\
  % \vskip -2cm
  \includegraphics[width=0.45\linewidth, angle=-90, trim={3cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-TiO2.eps}
\end{figure}

\begin{figure}[htbp]
  \centering
  {\bfseries \sffamily 2H-TiO$_{2}$}\\
  % \vskip -2cm
  \includegraphics[width=0.45\linewidth, angle=-90, trim={3cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-TiO2.eps}
\end{figure}

\begin{figure}[htbp]
  \centering
  {\bfseries \sffamily 2H-TiS$_{2}$}\\
  % \vskip -2cm
  \includegraphics[width=0.45\linewidth, angle=-90, trim={3cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-TiS2.eps}
\end{figure}

\begin{figure}[htbp]
  \centering
  {\bfseries \sffamily 2H-TiSe$_{2}$}\\
  % \vskip -2cm
  \includegraphics[width=0.45\linewidth, angle=-90, trim={3cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-TiSe2.eps}
\end{figure}

\begin{figure}[htbp]
  \centering
  {\bfseries \sffamily 1T-ZrO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-ZrO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-ZrS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-ZrS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-ZrSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-ZrSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-ZrO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-ZrO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-ZrSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-ZrSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-ZrTe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-ZrTe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-HfO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-HfO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-HfS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-HfS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-HfSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-HfSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-HfO4$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-HfO2.eps}
\end{figure}

% \begin{figure}[htbp]
% \centering
% {\bfseries \sffamily 2H-HfS$_{2}$}\\
% \includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-HfS2.eps}
% \end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-HfSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-HfSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-HfTe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-HfTe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-GeO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-GeO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-GeS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-GeS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-GeO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-GeO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-SnO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-SnO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-SnS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-SnS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-SnSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-SnSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-SnO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-SnO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-SnS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-SnS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-PbO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-PbO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily BN}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/hBN.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily C$_{2}$F$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/FGraph.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily P$_{4}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/Phosphorene.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily C$_{2}$H$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/Graphane.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-NiO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-NiO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-PdO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-PdO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-PdS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-PdS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-PtO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-PtO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-PtS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-PtS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-PdSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-PdSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-NiS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-NiS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 1T-PtSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/1T-PtSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily Ga$_{2}$Se$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/GaSe.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily Ga$_{2}$S$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/GaS.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily CdI$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/CdI2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-MoS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-MoS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-WS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-WS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-WSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-WSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-WO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-WO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-MoO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-MoO2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-MoTe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-MoTe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-WTe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-WTe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-CrS$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-CrS2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-CrSe$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-CrSe2.eps}
\end{figure}

\begin{figure}[htbp]
\centering
{\bfseries \sffamily 2H-CrO$_{2}$}\\
\includegraphics[width=0.45\linewidth, angle=-90, trim={2.9cm, 0cm, 2cm, 0cm}, clip]{img/SI_figs/BS/2H-CrO2.eps}
\end{figure}

% \end{document}