### 2018-10-01 DOS_figs.tex

\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-TiO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-TiO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-TiS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-TiSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-ZrO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-ZrS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-ZrSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-ZrO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-ZrSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-ZrTe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-HfO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-HfS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-HfSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-HfO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-HfS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-HfSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-HfTe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-GeO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-GeS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-GeO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-SnO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-SnS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-SnSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-SnO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-SnS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-PbO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/hBN-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/FGraph-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/Phosphorene-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/Graphane-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-NiO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-PdO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-PdS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-PtO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-PtS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-PdSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-NiS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/1T-PtSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/GaSe-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/GaS-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/CdI2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-MoS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-WS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-WSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-WO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-MoO2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-MoTe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-WTe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-CrS2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-CrSe2-DOS.pdf}
\end{center}
\begin{center}
\includegraphics[width=.9\linewidth]{img/SI_figs/2H-CrO2-DOS.pdf}
\end{center}