### 2018-11-15 SI.tex
Add analytical solution to anisotropy

% Created 2017-10-05 Thu 16:11
% Intended LaTeX compiler: pdflatex
\documentclass[manuscript=suppinfo,email=true,hyperref=true,keywords=false]{achemso}
\usepackage{times}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage{ltablex}
\usepackage{footnote}
\usepackage[colorinlistoftodos,prependcaption,textsize=small]{todonotes}

\usepackage{xr}
\externaldocument[main-]{paper}

\SectionNumbersOn{}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\theequation}{S\arabic{equation}}

\author{Tian Tian}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\altaffiliation{T. T. and D. S. contributed equally to this work}
\author{Declan Scullion}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}
\altaffiliation{T. T. and D. S. contributed equally to this work}
\author{Dale Hughes}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}
\author{Lu Hua Li}
\affiliation{Institute for Frontier Materials, Deakin University, Waurn Ponds, Victoria, Australia}
\author{Chih-Jen Shih}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\author{Jonathan Coleman}
\affiliation{School of Physics, Centre for Research on Adaptive Nanostructures and Nanodevices (CRANN) and Advanced Materials and BioEngineering Research (AMBER), Trinity College Dublin, Dublin 2, Ireland.}
\author{Manish Chhowalla}
% \affiliation{Materials Science and Engineering, Rutgers University, 607 Taylor Road, Piscataway, New Jersey 08854, USA.}
\affiliation{Department of Materials Science \& Metallurgy, University of Cambridge, CB3 0FS, United Kindom}
\author{Elton J. G. Santos}
\email{e.santos@qub.ac.uk}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}

\date{}

\title{Universal Understanding of the Dielectric Nature of Two-Dimensional Materials}
\begin{document}

\newpage{}
\section{Further Analysis of the 2D Polarizability Model}
\label{sec:2D-pol}

In this section, we discuss further topics about the 2D polarizability
model proposed in the main text.

\subsection{Profile of induced dipoles of 2D material}
\label{sec:dipole-plot}
Here we show in detail the
$\Delta \rho=\rho(\boldsymbol{E}) - \rho(\boldsymbol{E}=0)$ profile of
the 2H-MoS$_{2}$ slab in main text Figure \ref{main-fig-1}a. The
density $\rho$ is calculated as the average value in the
\textit{xy}-plane, and the value is shown in Figure
\ref{fig:rho-profile}. As can be seen the induced charges on the
MoS$_{2}$ layer only extends to a width of $\sim{}$12 \AA{}. When the
SL size $L \gg$12 \AA{} as in the first principle calculations done in
the main text, the induced dipoles from the periodic images do not
interact, and thus giving converged values of $\alpha_{\mathrm{2D}}$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/SI-dipole-MoS2.pdf}
  \caption{Induced dipole profile of a 2H-MoS$_{2}$ slab inside the
    superlattice, corresponding to main text Figure
    \ref{main-fig-1}a. \textbf{a}. Scheme of the MoS$_{2}$ layer
    inside the SL separated by vacuum layer. The dielectric constant
    $\varepsilon_{\mathrm{SL}}$ is combined effect of the 2D
    polarizability $\alpha_{\mathrm{2D}}$ and the vacuum
    layer. \textbf{b}. $\Delta \rho$ as a function of $z$ around the
    MoS$_{2}$ layer, green and red parts corresponding to negative and
    positive induced charges. The external electric field
    $E_{\mathrm{ext}}$ is 0.01 eV/\AA{}.}
  \label{fig:rho-profile}
\end{figure}

\subsection{Hypothetical 2D ``dielectric constant'' rescaled from 3D dielectric constant}
\label{sec:2D-3D-rescale}
From literatures, it is often desired to extract the ``dielectric
constant'' of a 2D material from the 3D macroscopic dielectric
constant of the superlattice, where the layers are periodic along of
x,y directions but separated by a vacuum layer between its images
\cite{Matthes_2016,Laturia_2018}. As discussed in the main text, the
definition of the 2D ``dielectric constant'' (in particular, its
in-plane components) is questionable, due to the locality of
electrostatic screening of the 2D materials
\cite{Cudazzo_2010_screen2D,Cudazzo_2011_screening_2D}. Here we show
that, in addition to the ambiguous definition of the 2D dielectric
constant, such approach requires the precise determination of the 2D
layer thickness $\delta_{\mathrm{2D}}$, which makes it impractical.

In brief, the hypothetical 2D dielectric constant
$\varepsilon$ comes from the planar limit of the
Maxwell-Garnett effective medium theory \cite{Markel_2016}, and can be
expressed by means of the macroscopic dielectric constant
$\varepsilon_{\mathrm{SL}}$ of the superlattice:

\begin{eqnarray}
  \label{eq:MG-effect-1}
  &\varepsilon_{\mathrm{SL}}^{\parallel} &= f \varepsilon_{\mathrm{2D}}^{\parallel} + (1 - f)\\
  \label{eq:MG-effect-2}
  &(\varepsilon_{\mathrm{SL}}^{\perp})^{-1} &= f /\varepsilon_{\mathrm{2D}}^{\perp} + (1-f)
\end{eqnarray}
where $f=\delta/L$ is the volume fraction of the 2D
material in the superlattice. The major issue when using such rescale
relations comes from the determination of $\delta_{\mathrm{2D}}$. To
eliminate the modeling error caused by the \textit{a priori} parameter
selection of $\delta$, we perform the calculation of
$\varepsilon_{\mathrm{SL}}$ of group 6 TMDCs against different $L$,
and use least-square fitting to extract both
$\varepsilon$ and $\delta$, as shown in
Figure \ref{fig:rescale-prb}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.85\linewidth]{img/SI-rescale-PRB.pdf}
  \caption{ Calculated (blue dots) and fitted
    (orange broken lines) $\varepsilon_{\mathrm{SL}}$ as function of
    $L$ for the group 6 TMDCs:
    \textbf{a}. MoS$_{2}$. \textbf{b}. MoSe$_{2}$. \textbf{c}.
    MoTe$_{2}$. \textbf{d}. WS$_{2}$. \textbf{e}. WSe$_{2}$. \textbf{f}.
    WTe$_{2}$. The extracted values of $\varepsilon_{\mathrm{2D}}$ and
    $\delta$ are shown in each subfigure.}
  \label{fig:rescale-prb}
\end{figure}

As can be seen from the fitted results in Figure
\ref{fig:rescale-prb}, the values of $\delta_{\mathrm{2D}}$ extracted
from both $\varepsilon_{\mathrm{SL}}^{\parallel}$ and
$\varepsilon_{\mathrm{SL}}^{\perp}$ are close when $L> 15$
{\AA}. Notably, the $\delta_{\mathrm{2D}}$ values are generally 10\%
smaller than the interlayer distance in corresponding bulk materials
$L_{\mathrm{bulk}}$ , as shown in Table \ref{tab:delta-L-DFt}. On the
other hand, the extracted $\delta_{\mathrm{2D}}$ values are closer to
the covalent thickness $\delta_{\mathrm{2D}}^{\mathrm{cov}}$ as
described in the main text, with a difference generally smaller than
5\%. Our results indicate that the conventional estimation of the 2D
layer thickness by its bulk interlayer distance
\cite{Matthes_2016,Laturia_2018}, will always lead to
overestimation. On the contrary, the out-of-plane polarizability
$\alpha_{\mathrm{2D}}^{\perp}$ correctly captures the thickness nature
of 2D materials.

\begin{table}[htbp]
  \centering
  \begin{tabular}[htbp]{lccccc}
  \hline{}
  Material & $\delta_{\mathrm{2D}}$ from $\varepsilon_{\mathrm{SL}}^{\parallel}$ ({\AA}) & $\delta_{\mathrm{2D}}$ from $\varepsilon_{\mathrm{SL}}^{\perp}$ ({\AA})& $L_{\mathrm{bulk}}$ ({\AA}) & $\delta_{\mathrm{2D}}^{\mathrm{cov}}$ ({\AA}) & $\alpha_{\mathrm{2D}}^{\perp}/\varepsilon_{0}$ ({\AA})\\
  \hline{}
  2H-MoS$_{2}$ & 5.76 & 5.49 & 6.15 & 5.22 & 4.98\\
  2H-MoSe$_{2}$ & 5.98 & 5.92 & 6.46 &  5.73 & 5.60\\
  2H-MoTe$_{2}$ & 6.43 & 6.85 & 6.98 & 6.37 & 6.12\\
  2H-WS$_{2}$ & 5.63 & 5.49 & 6.15 & 5.20 & 5.00\\
  2H-WSe$_{2}$ & 5.84 & 5.92 & 6.49 & 5.75 & 5.42\\
  2H-WTe$_{2}$ & 6.32 & 6.58 & 7.06 & 6.38 & 6.33\\
  \hline{}
\end{tabular}

\caption{Calculated $\delta_{\mathrm{2D}}$ from
  $\varepsilon_{\mathrm{SL}}^{\parallel}$ and
  $\varepsilon_{\mathrm{SL}}^{\perp}$, compared with the interlayer
  distance of corresponding bulk material $L_{\mathrm{SL}}$, the
  covalent thickness $\delta_{\mathrm{2D}}^{\mathrm{cov}}$, and
  $\alpha_{\mathrm{2D}}^{\perp}/\varepsilon_{0}$ for 2H TMDC materials.}
\label{tab:delta-L-DFt}
\end{table}

Another drawback of the 2D dielectric constant approach is the
overestimation of the out-of-plane dielectric response. As can be seen
in Figure \ref{fig:rescale-prb}, the extracted
$\varepsilon_{\mathrm{2D}}^{\perp}$ values for the TMDCs studied
are comparable or even larger than
$\varepsilon_{\mathrm{2D}}^{\parallel}$, which does not agree with the
physical picture that electrostatic screening of 2D materials are much
smaller perpendicular to the 2D plane. In fact, combining
Eq. \ref{eq:MG-effect-1} and the definition of $\alpha_{\mathrm{2D}}^{\perp}$, we
have:
\begin{equation}
  \label{eq:eps-alpha-perp}
  \frac{\alpha_{\mathrm{2D}}^{\perp}}{\varepsilon_{0}} = \delta_{\mathrm{2D}}(1 - (\varepsilon_{\mathrm{2D}}^{\perp})^{-1})
\end{equation}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{img/SI-eps-alpha-error.pdf}
  \caption{Calculated $\varepsilon_{\mathrm{2D}}^{\perp}$ value as a
    function of the difference between $\delta_{\mathrm{2D}}$ and
    $\alpha_{\mathrm{2D}}^{\perp}/\varepsilon_{0}$. A small change of
    $\alpha_{\mathrm{2D}}^{\perp}/\varepsilon_{0}$ chosen may lead to divergence of
    the $\varepsilon_{\mathrm{2D}}^{\perp}$ or even negative values,
    which is apparently unphysical}
  \label{fig:eps-alpha-error}
\end{figure}

which indicate that the characteristic length
$\alpha_{\mathrm{2D}}^{\perp}/\varepsilon_{0}$, is very close but slightly smaller
than $\delta_{\mathrm{2D}}$ estimated by the effective medium theory,
if $\varepsilon^{\perp}_{\mathrm{2D}} \gg 1$. Moreover, from
Eq. \ref{eq:eps-alpha-perp}, when $\delta_{\mathrm{2D}}$ and
$\alpha_{\mathrm{2D}}^{\perp}/\varepsilon_{0}$ are close, slight change of the
$\delta_{\mathrm{2D}}$ chosen may lead to divergence of
$\varepsilon_{\mathrm{2D}}^{\perp}$, as shown in Figure
\ref{fig:eps-alpha-error}. Therefore cautions must to taken when
treating the dielectric response of the 2D material using effective
medium theory. In comparison, the 2D polarizability does not require
the guess of the thickness, and therefore is the true descriptor of
the 2D dielectric nature.

To conclude, based on theoretical and technical considerations, there are several
advantages of using the polarizability $\alpha$ for describing the dielectric
nature of 2D materials, including
\begin{enumerate}
\item $\alpha_{\mathrm{2D}}$ can be used to describe both the local and macroscopic dielectric properties, while $\epsilon_{\mathrm{2D}}$ can't.
\item Calculating $\alpha_{\mathrm{2D}}$ only requires to calculate the dielectric response at single superlattice length, while $\epsilon_{\mathrm{2D}}$ requires calculation with varied superlattice length.
\item $\alpha_{\mathrm{2D}}$ correctly respresents the screening length of 2D material, while $\epsilon_{\mathrm{2D}}$ does not.
\item $\alpha_{\mathrm{2D}}$ correctly represents the different degree of screening in/out-of-plane, while $\epsilon_{\mathrm{2D}}$ does not.

\item The value of $\epsilon_{\mathrm{2D}}$ hugely depends on the
  choice of the thickness of 2D material, while such information is
  intrinsically embedded in $\alpha_{\mathrm{2D}}$.
\end{enumerate}

\subsection{Dependence of $\alpha_{\mathrm{2D}}$ on bandgap}
\label{sec:pol-2D-Eg}

In this section we further look into the bandgap dependency of the 2D
polarizability. Figure \ref{fig:SI-raw-HSE} shows the raw data of
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
as functions of $E_{\mathrm{g}}$ of the 2D materials investigated. We
observe that $\alpha_{\mathrm{2D}}^{\parallel}$ can be approximated by
a reciprocal function of $E_{\mathrm{g}}$, that
$\alpha_{\mathrm{2D}}^{\parallel}\sim{}
7.295(E_{\mathrm{g}})^{-1}$. On the other hand, the plot of
$\alpha_{\mathrm{2D}}^{\perp}$ against $E_{\mathrm{g}}$ shows no
apparent correlation.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-alpha_EG_raw_HSE.pdf}
  \caption{$E_{\mathrm{g}}$-dependence of \textbf{a} $\alpha_{\mathrm{2D}}^{\parallel}$ and
    \textbf{b} $\alpha_{\mathrm{2D}}^{\perp}$ for the 2D materials investigated here.}
  \label{fig:SI-raw-HSE}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{img/SI-compare_alpha_Eg.pdf}
  \caption{Relation between $1/\alpha_{\mathrm{2D}}^{\parallel}$ and various choices
    of $E_{\mathrm{g}}$: minimal gap from HSE06 (blue), minimal gap
    from PBE (orange) and direct gap from PBE (green). The linear
    regression results are shown as broken lines.}
  \label{fig:alpha-Eg-diff}
\end{figure}

We also investigate the relation of 2D polarizabilities with
difference choices of \textit{ab initio} bandgaps. It is widely
accepted that the PBE exchange correlation, tends to underestimate the
bangap \cite{Heyd_2005,Kumar_2016_PRB,Kumar_2016_jpcc}.  Indeed,
changing the choice of $E_{\mathrm{g}}$ yields different regression
relation with $1/\alpha_{\mathrm{2D}}^{\parallel}$, as shown in Figure
\ref{fig:alpha-Eg-diff}. We see that due to the underestimation of PBE
bandgap, the slope of linear regression is larger than that from
HSE-bandgap. We also observe that the $1/\alpha-E_{\mathrm{g}}$
relation is better presented by using the minimal HSE bandga than the
minimal PBE bandgap, due to higher regression $R^{2}$ coefficient of
the former. We note that the higher $R^{2}$ coefficient observed using
the direct PBE bandgap than the minimal PBE bandgap may be solely
caused by the fact that the direct bandgap of 2D materials on the PBE
level is closer to the HSE bandgap. From the random phase
approximation theory of dielectric response, the polarizability is
contributed by all possible transition between valence and conduction
bands, with the minimal bandgap as the least possible transition. In
this sense, $\alpha_{\mathrm{2D}}^{\parallel}$ is mostly like to be associated with
the minimal, not direct bandgap, as also observed in the original Moss
relation. We also examine the validity of such statement based on the
analysis of the Computational 2D Materials Database (C2DB)
\cite{Haastrup_2018}, as will be discussed in the following sections.

\section{Validation on the C2DB 2D materials database}
\subsection{Validation of the universal description of 2D polarizabilities}
\label{sec:gpaw-1}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{img/SI-gpaw-data.pdf}
  \caption{Validation the linear relation between \textbf{a}
    $1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$ and \textbf{b}
    $\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ on the C2DB 2D material
    database, corresponding to Figure 3b and 3c in main text.}
  \label{fig:gpaw-alpha-relation}
\end{figure}

Due to the high computation complexity of dielectric response
calculations on the HSE level, our 2D material database is limited to
about 50 types of materials. It is desirable to validate our proposed
relations on a even larger scale database. We select over 230
semiconducting 2D materials from the C2DB database with a GW bandgap
larger than 0.05 eV and extracted the 2D polarizabilities calculated
on the PBE level. The propsed linear relations between
$1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$ and
$\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ are also valid,
as shown in Figure \ref{fig:gpaw-alpha-relation}. Excellent linear
correlation is observed in both cases, with the $R^{2}$ coefficient
larger than 0.9, indicating the existence of the universal description
of 2D dielectric nature through the proposed relations with the 2D
polarizabilities. We note that the slope of linear regression is
slightly different from the our proposed relation from the dielectric
response on HSE06 level, due to the fact that the macroscopic
dielectric constant $\varepsilon_{\mathrm{SL}}$ calculated on the PBE
level is about slightly difference. Despite the minor deviation of the
slope of linear regression, the analysis on the C2DB database supports
our understanding of the universal dielectric nature of 2D materials.

\subsection{Choice of bandgap}
\label{sec:gpaw-2}

Next we investigate the influence of choice of $E_{\mathrm{g}}$ on the
regression of $\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}$ relation. Figure
\ref{fig:SI-gpaw-alpha-Eg-all} shows $1/\alpha_{\mathrm{2D}}^{\parallel}$ from the
C2DB database as a function of minimal and direct bandgap calculated
on PBE, HSE06 and GW levels. We observe,
although the regression $R^{2}$ coefficient in all cases are around
0.9, the $\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}$ is better described using
the HSE and GW bandgaps than the PBE bandgaps. On the other hand,
using indirect or minimal bandgaps on the same level gives almost
identical regression slope. The observations are in good agreement
with our calculations on the HSE level discussed in Section
\ref{sec:pol-2D-Eg}. In combination with the physical contribution of
$E_{\mathrm{g}}$ to the dielectric screening, we conclude that the
minimal bandgap should be used for quantitative prediction of the
in-plane 2D polarizability. The prediction is greatly improved when
more accurate theory level for bandgap is used (for instance, HSE and GW).

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/SI-gpaw-alpha_Eg_all.pdf}
  \caption{$\alpha_{\mathrm{2D}}^{\parallel}$ as funtion of minimal and direct
    $E_{\mathrm{g}}$ calculated on different theoretical levels:
    \textbf{a} PBE, \textbf{b} HSE and \textbf{c} GW of the C2DB
    database.}
  \label{fig:SI-gpaw-alpha-Eg-all}
\end{figure}

For 2D materials, the exciton effect plays an important role in
determining the experimentally accessible bandgap
\cite{Arnaud_2006_exc_hBN,Pulci_2014,Ramasubramaniam_2012,Chernikov_2014_EB_MoS2_2D3D}. The
experimentally observed optical bandgap
$E_{\mathrm{g}}^{\mathrm{opt}}$, is usually lower than the direct
bandgap from bandstructure $E_{\mathrm{g}}^{\mathrm{dir}}$ by the
exciton binding energy $E_{\mathrm{b}}$, which is at the 10$^{-1}$ to
10$^{1}$ eV for different 2D materials due to the attenuated
dielectric screening. Next we examine the relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{\mathrm{opt}}$
from the C2DB database, with
$E_{\mathrm{g}}^{\mathrm{opt}}=E_{\mathrm{g}}^{\mathrm{dir,QP}}-E_{\mathrm{b}}^{\mathrm{BSE}}$,
where $E_{\mathrm{g}}^{\mathrm{dir,QP}}$ is the direct quasi-particle
bandgap calculated using G0W0 method and
$E_{\mathrm{b}}^{\mathrm{BSE}}$ is the exciton binding energy
calculated using the Bethe-Salpethe equation. Figure \ref{fig:opt}
shows $(4\pi\varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ as a
function of $E_{\mathrm{g}}^{\mathrm{opt}}$, with a linear regression
slope of 0.154 and $R^{2}$ of 0.84, similar to the relation between
$(4\pi \varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ and
$E_{\mathrm{g}}$ (from HSE06 level, see Figures \ref{fig:SI-raw-HSE}
and \ref{fig:SI-gpaw-alpha-Eg-all}). The roughly linear correlation
between $(4\pi \varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ and
$E_{\mathrm{g}}^{\mathrm{opt}}$ is not coincidence: in fact,
theoretical analysis shows that the binding energy $E_{\mathrm{b}}$ is
proportional to the direct bandgap $E_{\mathrm{g}}$
\cite{Jiang_2017_Eg_Eb}, taking into account that
$(\alpha_{\mathrm{2D}}^{\parallel})^{-1} \propto E_{\mathrm{g}}$, we
rationalize that
$(\alpha_{\mathrm{2D}}^{\parallel})^{-1} \propto
E_{\mathrm{g}}^{\mathrm{opt}}=E_{\mathrm{g}}^{\mathrm{dir}}-E_{\mathrm{b}}$. The
slightly smaller linearity than the 2D Moss-like relation is caused
from multiple approximations used. Nevertheless we show that
$(4\pi\varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ can be
equivalently predicted using the experimentally accessible optical
bandgap.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.55\linewidth]{img/SI-alphax-Eg-opt.pdf}
  \caption{$(4\pi\varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ as a function of
    $E_{\mathrm{g}}^{\mathrm{opt}}$ from the C2DB database.}
  \label{fig:opt}
\end{figure}

\subsection{Relation between 2D polarizabilities and other physical quantities}
\label{sec:gpaw-3}

The relatively large size of the C2DB 2D materials database allows us
to examine the relation between 2D polarizabilities and other physical
quantities. We choose the following quantities for comparison,
corresponding to Figures \ref{fig:gpaw-2D-quantities-1} to
\ref{fig:gpaw-2D-quantities-3}:
\begin{enumerate}
\item The effective carrier mass for electron $m_{e}^{*}$ and hole $m_{h}^{*}$

\item The quantum capacitance at the conduction band edge
  $C_{\mathrm{Q}}^{\mathrm{C}}$ and valence band edge
  ($C_{\mathrm{Q}}^{\mathrm{V}}$).

\item The total atomic polarizabilities per area $\alpha_{\mathrm{2D}}^{\mathrm{sum}}$.
\end{enumerate}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.75\linewidth]{img/SI-gpaw-alpha-emass.pdf}
  \caption{Relation between 2D polarizabilities and the effective
    carrier mass from the C2DB 2D material
    database. \textbf{a}. $\alpha_{\mathrm{2D}}^{\parallel}$ as a function of the
    electron mass $m_{e}^{*}$.  \textbf{b}. $\alpha_{\mathrm{2D}}^{\parallel}$ as a
    function of the hole mass
    $m_{h}^{*}$. \textbf{c}. $\alpha_{\mathrm{2D}}^{\perp}$ as a function of the
    electron mass $m_{e}^{*}$.  \textbf{d}. $\alpha_{\mathrm{2D}}^{\perp}$ as a
    function of the hole mass $m_{h}^{*}$. No apparent correlation
    between the 2D polarizabilities and the effective carrier masses
    is observed.}
  \label{fig:gpaw-2D-quantities-1}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{img/SI-quantum-capacitance.pdf}
  \caption{Relation between the 2D polarizabilities with the quantum
    capacitance. \textbf{a} $\alpha_{\mathrm{2D}}^{\parallel}$ (top) and
    $\alpha_{\mathrm{2D}}^{\perp}$ (bottom) as functions of the quantum capacitance
    of the conduction band edge,
    $C_{\mathrm{Q}}^{\mathrm{C}}$. \textbf{b} $\alpha_{\mathrm{2D}}^{\parallel}$
    (top) and $\alpha_{\mathrm{2D}}^{\perp}$ (bottom) as functions of the quantum
    capacitance of the valence band edge,
    $C_{\mathrm{Q}}^{\mathrm{V}}$. Similar to the case of effective
    carrier mass, no apparent correlation between 2D polarizabilities
    and the quantum capacitance can be found.}
  \label{fig:gpaw-2D-quantities-2}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.75\linewidth]{img/SI-gpaw-alpha-atomic-polarizability.pdf}
  \caption{Relation between the 2D polarizabilities
    (\textbf{a}. $\alpha_{\mathrm{2D}}^{\parallel}$ and
    \textbf{b}. $\alpha_{\mathrm{2D}}^{\perp}$) with the total atomic polarizability per area.}
  \label{fig:gpaw-2D-quantities-3}
\end{figure}

The quantum capacitance
$C_{\mathrm{Q}}(E)$ at certain energy level $E$ is calculated using
the relation $C_{\mathrm{Q}}(E)=\mathrm{DOS}(E)e^{2}$, where
$\mathrm{DOS}(E)$ is the density of states at the conduction or valence band edge (averaged by cell
area). The DOS value is calculated at the energy level with a charge
cutoff such that
$|n_{\mathrm{2D}}(E)| = 5 \times 10^{13}\ \mathrm{cm}^{-2}$, calculated by
the relation of accumulated charge $n_{\mathrm{2D}}(E)$ at CB or VB:
\begin{equation}
  \label{eq:CQ-method}
  |n_{\mathrm{2D}}(E)| = \left|\int_{E_{\mathrm{BE}}}^{E} \mathrm{DOS}(E') dE' \right|
\end{equation}
where $E_{\mathrm{BE}}$ is the energy of the CB or VB band edge.

The total polarizability $\alpha_{\mathrm{2D}}^{\mathrm{sum}}$ is calculated by the
summation of the atomic polarizabilities $\alpha_{\mathrm{2D}}^{\mathrm{atom}}$
\cite{Gould_2016_jctc} of individual atoms per area $A$, such that:
\begin{equation}
  \label{eq:atom-polar}
  \alpha_{\mathrm{2D}}^{\mathrm{sum}} = \frac{\sum_{i} \alpha^{\mathrm{atom}}_{i}}{A}
\end{equation}

From Figures \ref{fig:gpaw-2D-quantities-1} to
\ref{fig:gpaw-2D-quantities-3} we can see that none of the above
quantities have apparent relation with the 2D polarizabilities, as
compared with the bandgap and covalent thickness proposed in the main
text.

\section{Theoretical Analysis}
\label{sec:theory}
\subsection{Polarizability-Based Theoretical Model}
\label{ssec:theory-1}
The universal relations for $\alpha_{\mathrm{2D}}^{\parallel}$ and
$\alpha_{\mathrm{2D}}^{\perp}$ revealed by main text
Eqs. \ref{main-eq:2D-Moss-para} and \ref{main-eq:2D-Moss-perp} are not
coincidence. Combining recent theoretical findings of the linear
relation between exciton binding energy $E_{\mathrm{b}}$ and
$E_{\mathrm{g}}$ of 2D materials
\cite{Choi_linear_2015,Olsen_2016_hydrogen,Jiang_2017_Eg_Eb}, and the
fact that the $E_{\mathrm{b}}$ is roughly inversely proportional to
$\alpha_{\mathrm{2D}}^{\parallel}$ \cite{Pulci_2014}, it is reasonable
to have a general Moss-like relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{-1}$. Moreover,
the bandgap-independent relation of 2D $\alpha_{\mathrm{2D}}^{\perp}$
resembles molecular polarizabilities of conjugate molecules
\cite{Davies_1952}, fullerenes \cite{Sabirov_2014} and carbon
nanotubes \cite{Benedict_1995}, which are also shown to be
geometry-dependent.

In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wavefunction of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components, similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$, such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, c, v are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{main-eq:2D-Moss-para} and
\ref{main-eq:2D-Moss-perp}.

\subsubsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-para}}
\label{sssec:theory-1}

In this section we show how main text Eq. \ref{main-eq:2D-Moss-para}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes the independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{v}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Present the summation of k-space into integral, and further combining
with the Bloch wave representation and the
$\mathbf{k} \cdot \mathbf{p}$ theory\cite{Jiang_2017_Eg_Eb}, we arrive
at Eq. \ref{main-eq:2D-Moss-para}. Note that from
Ref. \citenum{Jiang_2017_Eg_Eb}, the coefficient between
$\alpha^{\parallel}_{\mathrm{2D}}$ and $E_{\mathrm{g}}$ is given by:
$\alpha_{\mathrm{2D}}^{\parallel} = N e^{2}/(2\pi E_{\mathrm{g}})$,
where $N$ is the degeneracy associated with the minimal bandgap. Such
relation predicts the linear correlation coefficient between
$(4\pi \varepsilon_{0})/\alpha^{\parallel}_{\mathrm{2D}}$ (measured in
\AA{}$^{-1}$) and $E_{\mathrm{g}}$ (measured in eV) as
$8 \pi^{2} \varepsilon_{0} \AA{} / (e N) \approx 0.436 / N$. From main
text Figure \ref{main-fig-3}c, we have
$(4 \pi \varepsilon_{0}\cdot \text{\AA{}})/\alpha_{\mathrm{2D}}^{\parallel}
\sim{} 0.183 E_{\mathrm{g}}/ \mathrm{eV}$, corresponding to $N$
between 2 and 3, which is a reasonable result for the 2D materials
studied.

\subsubsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-perp}}
\label{ssec:theory-2}
For main text Eq. \ref{main-eq:2D-Moss-perp}, treating the in-plane
wavefunctions as plane wave with form
$\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the matrix
element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z-2}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at:
\begin{equation}
\label{eq:alpha_perp_RPA}
\alpha_{\mathrm{2D}}^{\perp} = \frac{2e^{2}}{(2 \pi) ^{2}} \int d^{2}\mathbf{k}
\sum_{\mathrm{c, v}}
\frac{|<\psi_{\mathrm{v}}(\mathbf{k})|z|\psi_{\mathrm{c}}(\mathbf{k})>|^{2}}
{E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
The formalism is slightly different from Eq.\ref{eq:alpha_para_RPA}.

The out-of-plane wavefunction $\psi^{\perp}(z)$ is the solution to the Schr\"{o}dinger equation with
Hamiltonian $\mathcal{H} = -\hbar^{2} \nabla^{2}/2m_{e} + V(z)$, where
$\hbar$ is the reduced Planck constant, $m_{e}$ is electron mass and
$V(z)$ is the confined Coulombic potential along the z-direction
created by the nuclei
\cite{davies_physics_1997,ihn_semiconductor_2009}. Although the exact
form for $\psi^{\perp}$ depends on the exact distribution of $V(z)$,
without loss of generality we can assume the electrons are confined in
a potential well of width $\delta$, which is the typical treatment for
semiconductor QWs \cite{ihn_semiconductor_2009}. The allowed bound
states inside the confined region generally has wave vector
$k_{z} \propto n \pi / \delta$. With the total energy
$E_{n}(\mathbf{k}) = {\displaystyle \frac{\hbar^{2} (k_{x}^{2} +
    k_{y}^{2})}{2 m^{\parallel}} + \frac{\hbar^{2} n^{2} \pi^{2}}{2
    m^{\perp} \delta^{2}}}$, where $m^{\parallel}$ and $m^{\perp}$ are
the effective masses parallel and perpendicular to the 2D
plane. Therefore the denominator of Eq. 10 becomes independent of
$\mathbf{k}$, that
$E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k}) =
(n_{\mathrm{c}}^{2} - n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{\hbar^{2} \pi^{2}}{2 m^{\perp} \delta^{2}}}$. On the other
hand, the numerator
$<\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>$
is proportional to the confinement length $\delta$ (can be seen using
particle-in-box solution\cite{davies_physics_1997}). In combination,
the individual terms of the summation in the right hand of Eq. 5b is
independent of neither $E_{\mathrm{g}}$ nor \textbf{k}, proving that
$\alpha_{\mathrm{2D}}^{\perp}$ is independent of the band gap.

Finally we will show the $\alpha_{\mathrm{2D}}^{\perp}$ is related to the
confinement length $\delta$, or equivalent the thickness of the 2D
material. For simplicity we assume $\psi^{\parallel}$ is the solution
of a particle confined inside an infinite potential well of width
$\delta$ with the form $\psi \propto \sin(\frac{n\pi}{\delta})$. At 0
K, the only non-zero contribution for the transition between band
$n_{\mathrm{v}}$ and $n_{\mathrm{c}}$ corresponds with in-plane wave
vector $\mathbf{k}_{\rho}$ within the range of $k_{\mathrm{F}}^{\mathrm{v}}$ to
$k_{\mathrm{F}}^{\mathrm{c}}$, the Fermi vectors of each band. Therefore we have:
\begin{equation}
  \begin{aligned}
    \alpha_{\mathrm{2D}}^{\perp} &\propto \delta^{4} m^{\perp} \int
    d^{2}\mathbf{k}_{\rho}
    \frac{f(\psi_{\mathrm{v}}(\mathbf{k}_{\rho}))
      -f(\psi_{\mathrm{v}}(\mathbf{k}_{\rho})) }{n_{\mathrm{c}}^{2} -
      n_{\mathrm{v}}^{2}}\\
    &= \delta^{4} m^{\perp} \sum_{\mathrm{c, v}}\pi \frac{
      (k_{\mathrm{F}}^{\mathrm{c}})^{2} -
        (k_{\mathrm{F}}^{\mathrm{v}})^{2}}{n_{\mathrm{c}}^{2} -
        n_{\mathrm{v}}^{2}}
  \end{aligned}
\end{equation}
Since the Fermi wave vector for band \textit{i} follows
${\displaystyle
  \frac{\hbar^{2}(k_{\mathrm{F}}^{i})^{2}}{2m^{\parallel}}} +
{\displaystyle \frac{\hbar^{2} n_{i}^{2} \pi^{2}}{2m^{\perp}
    \delta^{2}}} = E_{\mathrm{F}}$, we have
$ (k_{\mathrm{F}}^{\mathrm{c}})^{2} -
(k_{\mathrm{F}}^{\mathrm{v}})^{2} \propto (n_{\mathrm{c}}^{2} -
n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{m^{\parallel}}{m^{\perp}}}$. Note that in the particle-in-box
model, the energy levels for a sub-nm QW will be at the order of 10
eV, which allows only few states inside the QW. Combine with the fact
that the strongest transition mostly comes from the lowest bands
\cite{davies_physics_1997}, $\alpha^{\perp}_{\mathrm{2D}}$ is shown to
be proportional to $\delta^{p}$, where the power index $p$ is 2. We
note the different power law comes from the choice of the $V(z)$. The
atomic polarizabilities of free particles inside a 1D infinite QW,
finite QW or harmonic QW all show dependency with $\delta^{4}$, where
$\delta$ is the length of confinement or effective harmonic length of
the QW \cite{Fowler_1984,Maize_2011}. Since the 2D polarizability is
the molecular polarizability divided by the area, such approach will
always yield $\alpha_{\mathrm{2D}}^{\perp} \propto \delta^{2}$. To
accurately model the $\alpha_{\mathrm{2D}}^{\perp}$ we need a better
model for the confinement in z-direction than simple square potential
well, which may be tedious. Nevertheless the simple model presented
here still captures the bandgap-indepent and thickness-related feature
of $\alpha_{\mathrm{2D}}^{\perp}$, supporting our findings in the main
text. In the next section we will provide a simple explanation for the $\alpha \propto \delta$ relation from fundamental electrostatics theory.

\subsubsection{Explanation to main text Eq. \ref{main-eq:2D-Moss-perp}
  from fundamental electrostatics}
\label{sssec:theory-3}

The dependency of $\alpha_{\mathrm{2D}}^{\perp}$ on the thickness
$\delta$ of a 2D material, can also be regarded using fundamental
electrostatic model. Consider the smallest repeating unit of the 2D
material with xy-plane area $A$, under small perturbation field $E$
along the z-direction.  Note that the surface bound charge
$\sigma_{\mathrm{b}}=n e /A$, where $n$ is the number of unit charges
contributes to the bound charges, comes only from the dipoles of the
outer-most atoms, since the induced charges from inner atoms are
cancel out (see Figure \ref{fig:classic-model}).
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7\linewidth]{img/SI-classic.pdf}
  \caption{Fundamental electrostatic model for the
    thickness-dependency of $\alpha_{\mathrm{2D}}^{\perp}$, using 2H-MoS2 as an
    example. Left: induced dipoles from individual atoms along the
    z-direction. The positive and negative induced charges from inner
    atoms cancel out. Right: simplied model for the thickness
    dependency of $\alpha_{\mathrm{2D}}^{\perp}$, where the surface dipole density
    $\boldsymbol{\mu}$ comes only from the outer-most atoms.}
  \label{fig:classic-model}
\end{figure}
From the definition of
$\alpha_{\mathrm{2D}}^{\perp}$, we have:
\begin{equation}
  \label{eq:alpha-classic}
  \alpha_{\mathrm{2D}}^{\perp} = \frac{\boldsymbol{u}_{z}}{E_{\mathrm{loc}} A}
  = \frac{(d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j}) \sigma_{\mathrm{b}}}{E}
\end{equation}
where $r_{\mathrm{cov}}^{i}$ and $r_{\mathrm{cov}}^{j}$ are the
covalent radii of the outer-most atoms, the characteristic length of
the dipole extension in z-direction, and $d_{\mathrm{max}}$ is the
z-distance between the nuclei of such atoms.  The field $E$
counterbalances the field from the surface bound charges and equals
$E = \sigma_{\mathrm{b}}/\varepsilon_{0}$. Therefore we have:
\begin{equation}
  \label{eq:alpha-classic-2}
  \alpha_{\mathrm{2D}}^{\perp} = (d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j})\varepsilon_{0}
                = \delta_{\mathrm{cov}} \varepsilon_{0}
\end{equation}
which explains the linear relation seen in main text Figure
\ref{main-fig-3}d. We can see that such simple model nicely captures
the thickness feature of $\alpha_{\mathrm{2D}}^{\perp}$, and
reproduces the right coefficient between $\delta_{\mathrm{cov}}$ and
$\alpha_{\mathrm{2D}}^{\perp}$. Finally, the schematic drawing of the
polarizability ellipsoid for a 2D material, is drawn in Figure
\ref{fig:ellipsoid}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{img/SI-ellipsoid.pdf}
  \caption{Polarizability ellipsoid of a 2D material, the long axis
    equals $\alpha_{\mathrm{2D}}^{\parallel}/2\varepsilon_{0}$ and the short axis
    equals  $\alpha_{\mathrm{2D}}^{\perp}/2\varepsilon_{0}$}
  \label{fig:ellipsoid}
\end{figure}

\subsection{Comparing 2D and 3D Moss relations}
\label{ssec:theory-2}

The 2D Moss-like relation $\alpha^{\parallel} \propto E_{g}^{-1}$ is
similar to the 3D Moss relation $\varepsilon \propto E_{g}^{-1/2}$,
with a different power law. Such difference in the power law can
indeed be explained by modern theory of dielectric properties. From
the 2D material to a bulk covalent semiconductor, the wavefunction
becomes periodic in all directions. Considering only one pair of
valence-conduction transition and uniform effective mass $m^{*}$,
extending the approach Eq. \ref{eq:alpha_perp_RPA} to the bulk
material, and using the Bloch presentation for wavefunctions in all
dimensions, we get \cite{Jiang_2017_Eg_Eb}:
\begin{equation}
  \begin{aligned}
    \label{eq:eps-bulk}
    \varepsilon_{\mathrm{bulk}} - 1 &\propto \int d^{3}\mathbf{k}
    {\displaystyle \frac{1}{(E_{\mathrm{g}} + {\displaystyle
          \frac{\hbar^{2} k^{2}}{m^{*}}})^{2}}}\\
    &= \int_{0}^{k_{\mathrm{BZ}}} {\displaystyle
      \frac{4 \pi k^{2}}{(E_{\mathrm{g}} + {\displaystyle \frac{\hbar^{2}
            k^{2}}{m^{*}}})^{2}}} dk
  \end{aligned}
\end{equation}
where $k_{\mathrm{BZ}}$ is the boundary for the Brillouin Zone. The
last step in Eq. \ref{eq:eps-bulk} assumes the integral within the
Brillouin Zone is equivalent to integral inside a sphere of
k-space. Let $\hbar^{2}/(2 m^{*})=\beta$, the integral becomes:
\begin{equation}
  \begin{aligned}
    \label{eq:integral-BZ-bulk}
    \varepsilon_{\mathrm{Bulk}} &\propto {\displaystyle \frac{2 \pi
        \mathrm{arctan}(\sqrt{\beta k^{2}/E_{\mathrm{g}}})}{\sqrt{E_{\mathrm{g}} \beta^{3}}}
        - \frac{2\pi k}{\beta(\beta +E_{\mathrm{g}}k^{2})}
      } \bigg\rvert_{0}^{k_{\mathrm{BZ}}}\\
      &\propto 1/\sqrt{E_{\mathrm{g}}}
  \end{aligned}
\end{equation}
when $\varepsilon_{\mathrm{Bulk}} \gg 1$. since generally
$\hbar^{2}k_{\mathrm{BZ}}^{2}/(2m^{*}) \gg
E_{\mathrm{g}}$ \cite{Finkenrath_1988}. The final result $\varepsilon_{\mathrm{Bulk}} \propto E_{\mathrm{g}}^{-1/2}$ recovers the original Moss relation for bulk semiconductors.

\subsection{Static 2D polarizability and 2D plasma frequency}
\label{ssec:omega-p}

A common approach for describing the bulk dielectric function of bulk
semiconductors is via the Lorentz oscillator model, where the
dielectric function is dominated by the plasma frequencies
$\omega_{\mathrm{3D}}^{\mathrm{p}}$ and bandgap $E_{\mathrm{g}}$ of
individual oscillators \cite{ketterson_physics_2016}. At zero optical
frequency and the static limit, the dielectric constant for single
oscillator is:
\begin{equation}
  \label{eq:eps-plas-3D}
  \varepsilon_{\mathrm{3D}} = 1 +
  \frac{\hbar^{2} (\omega_{\mathrm{3D}}^{\mathrm{p}})^{2}}{E_{\mathrm{g}}^{2}}
\end{equation}
where
$\omega_{\mathrm{3D}}^{\mathrm{p}} = {\displaystyle \sqrt{\frac{e^{2}
      n_{\mathrm{3D}}}{\varepsilon_{0} m_{e}}}}$, where
$n_{\mathrm{3D}}$ is the 3D number density of valence
electrons. Combine Eq. \ref{eq:eps-plas-3D} with Eq. 3 in main text,
we get:
\begin{equation}
  \begin{aligned}
  \label{eq:alpha-plas}
  \alpha_{\mathrm{2D}}^{\parallel} &= \frac{e^{2} n_{\mathrm{3D}} L}{m_{e} E_{\mathrm{g}}^{2}} \\
  &= \frac{e^{2} n_{\mathrm{2D}}}{m_{e} E_{\mathrm{g}}^{2}} \\
  &= \varepsilon_{0} \frac{\hbar^{2}
    (\omega_{\mathrm{2D}}^{\mathrm{p}})^{2}}{E_{\mathrm{g}}^{2}}
\end{aligned}
\end{equation}
where $n_{\mathrm{2D}} =n_{\mathrm{3D}} L$ is the 2D number density of
valence electrons and
$\omega_{\mathrm{2D}}^{p}=\omega_{\mathrm{3D}}^{p}\sqrt{L}$ is the 2D
plasma frequency at static limit \cite{Nazarov_2015_2D_3D}, as
discussed in the main text. Apparently $n_{\mathrm{2D}}$ and
$\omega_{\mathrm{2D}}^{p}$ defines the superlattice-independent 2D
quantity $\alpha_{\mathrm{2D}}^{\parallel}$, while its 3D counterpart
$\varepsilon_{\mathrm{3D}}$ is dependent on $L$. By defining the 2D
valence charge density
$\sigma_{\mathrm{2D}}^{\mathrm{v}}=n_{\mathrm{2D}}e$, we have also
calculated $\alpha_{\mathrm{2D}}^{\parallel}$ as a function of
$\sigma_{\mathrm{2D}}^{\mathrm{v}}/E_{\mathrm{g}}^{2}$ using the C2DB database, as shown in Figure \ref{fig:plasma}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.65\linewidth]{img/SI-lorentz-model.pdf}
  \caption{Calculated$\alpha_{\mathrm{2D}}^{\parallel}$ as a function of
    $\sigma_{\mathrm{2D}}^{\mathrm{v}}/E_{\mathrm{g}}^{2}$ using the
    C2DB database. The broken line shows the theoretical prediction
    from single-oscillator model.}
  \label{fig:plasma}
\end{figure}
It can be seen that, a large number of materials are close to the
theoretical value of
$\alpha_{\mathrm{2D}}^{\parallel} = \frac{e^{2} n_{\mathrm{3D}} L}{m_{e}
  E_{\mathrm{g}}^{2}}$ (broken line). However there are also many
violations to this simple relation, making such model not suitable for
quantitative prediction of the 2D dielectric nature, due to the
oversimplification of single Lorentz oscillator. Nevertheless, this
example shows excellently how the quantities in both dimensions are
related to each other.

\subsection{Comparing dielectric anisotropy}
\label{ssec:aniso}
\subsubsection{Choice of materials}
\label{sssec:aniso-1}
The dielectric anisotropy $\eta$ proposed in the main text is also
applied to other dimensions. Similar to the case of 2D and bulk
layered materials, $\epsilon$ is used to compare the anisotropy when
the material is periodic in all dimensions (bulk covalent materials),
while the polarizability $\alpha$ is used for reduced dimensional
materials.

The following types materials are chosen for comparison:
\begin{itemize}
\item Bulk covalent materials.  \\
  The list of materials and bandgap are chosen according to
  Ref. \citenum{sze_appendix_2006}, including IV-IV, III-V, II-VI, and
  IV-VI semiconductors. All these materials have isotropic dielectric
  properties.
\item Planar OSc \\
  The planar OScs include metal phthanocyanines, disk-like polycyclic
  aromatic hydrocarbon (PAHs), and benzene derivatives. The
  dimensionality of these materials are close to 2D materials due to
  their planar shape. The bandgap values (mostly at B3LYP density
  functional with 6-31G** basis sets) are extracted from the NIST
  Computational Chemistry Comparison and Benchmark Database
  (http://cccdb.nist.gov, Release 19, April 2018) and the
  polarizability values are obtained from
  Refs. \citenum{Miller_1990,Ramprasad_2006}.
\item Carbon nanotubes (CNT) \\
  Like 2D materials, CNTs are periodic along the 1D directions, and
  shoudl be treated in a similar way to get the polarizability
  proportional to [Length]$^{2}$ \cite{Benedict_1995}. Semiconducting
  zigzag and armchair CNTs are considered, with their electronic
  properties obtained from Refs
  \citenum{Benedict_1995,Matsuda_2010}. and dielectric properties
  obtained from
  Refs. \citenum{Benedict_1995,Jensen_2000,Brothers_2008}.
\item Linear OSc \\
  We choose the polyacenes (linear PAHs from benznene to nonacene) as
  a model system of linear OScs. The bandgaps are obtained from
  \citenum{Salzner_1998} and the polarizabilities are obtained from
  \citenum{Hinchliffe_2005}.
\item Fullerenes \\
  The bandgap of fullerens (C$_{n}$ where $n=60*m^{2}$ where
  $m=1\sim{}7$) are taken from \citenum{Lin_1994} and the
  polarizabilities are taken from \citenum{Martin_2008}. All these
  materials have isotropic polarizability due to the high symmetry.
\end{itemize}

\subsubsection{Explanation for the separation between 2D and 3D regimes in main text Figure \ref{main-fig:aniso}}
\label{sssec:aniso-2}
In this section we give an analytical explanation for the separation
between the dielectric anisotropy indices of 2D and their bulk
counterparts. From main text Eqs. \ref{main-eq:alpha-para-def},
\ref{main-eq:alpha-perp-def} and \ref{main-eq:anisotropy},
$\eta_{\mathrm{Bulk}}$ of a bulk layered material is expressed as:
\begin{equation}
  \label{eq:eta-bulk}
  \begin{aligned}[t]
    \eta_{\mathrm{Bulk}} &= \frac{\varepsilon_{\mathrm{Bulk}}^{\perp}}
    {\varepsilon_{\mathrm{Bulk}}^{\parallel}}\\
    &= \frac{1}{\left(1 + {\displaystyle
          \frac{\alpha_{\mathrm{Bulk}}^{\parallel}}{\varepsilon_{0}L_{\mathrm{Bulk}}}}\right)
      \left(1 - {\displaystyle \frac{\alpha_{\mathrm{Bulk}}^{\perp}}{\varepsilon_{0}L_{\mathrm{Bulk}}}} \right)}\\
    &= \frac{1}{ \left[1-\left( {\displaystyle
            \frac{\alpha_{\mathrm{Bulk}}^{\perp}}{\alpha_{\mathrm{Bulk}}^{\parallel}}}\right)
        \left(
          {\displaystyle \frac{\alpha_{\mathrm{Bulk}}^{\parallel}}{\varepsilon_{0}
            L_{\mathrm{Bulk}}}} \right) \right]
      \left[ 1 + \left(
          {\displaystyle \frac{\alpha_{\mathrm{Bulk}}^{\parallel}}{\varepsilon_{0}
            L_{\mathrm{Bulk}}}} \right)\right]
    }
    % &= \frac{\left({\displaystyle \frac{\varepsilon_{0}L_{\mathrm{Bulk}}}{\alpha^{\parallel}_{\mathrm{Bulk}}}}\right)^{2}}
    % {\left({\displaystyle \frac{\varepsilon_{0}L_{\mathrm{Bulk}}}{\alpha_{\mathrm{Bulk}}^{\parallel}_{\mathrm{2D}}}}\right)^{2}    
      % -\left({\displaystyle \frac{\alpha_{\mathrm{Bulk}}^{\perp}}{\alpha_{\mathrm{Bulk}}^{\parallel}}}\right)
      % \left({\displaystyle \frac{\varepsilon_{0} L_{\mathrm{Bulk}}}{\alpha_{\mathrm{Bulk}}^{\parallel}}}\right)
      % + \left({\displaystyle \frac{\varepsilon_{0} L_{\mathrm{Bulk}}}{\alpha_{\mathrm{Bulk}}^{\parallel}}}\right)
      % -\left({\displaystyle \frac{\alpha_{\mathrm{Bulk}}^{\perp}}{\alpha_{\mathrm{Bulk}}^{\parallel}}}\right)}
  \end{aligned}
\end{equation}
Name take the fact that
$\alpha_{\mathrm{Bulk}}^{\parallel} \approx
\alpha^{\parallel}_{\mathrm{2D}}$ (main text Figure
\ref{main-fig-4}b), we name
$\alpha_{\mathrm{Bulk}}^{\parallel} /\epsilon_{0}L_{\mathrm{Bulk}}
\approx \alpha_{\mathrm{2D}}^{\parallel}/
\epsilon_{0}L_{\mathrm{Bulk}}$ as $\gamma$. Furthermore we name
$\alpha_{\mathrm{Bulk}}^{\perp}/\alpha_{\mathrm{Bulk}}^{\parallel}$ as
$\hat{\eta_{\mathrm{2D}}}$, that
$\hat{\eta}_{\mathrm{2D}} =
\alpha_{\mathrm{Bulk}}^{\perp}/\alpha_{\mathrm{Bulk}}^{\parallel} \geq
\alpha_{\mathrm{2D}}^{\perp}/\alpha_{\mathrm{2D}}^{\parallel}=\eta_{\mathrm{2D}}$,
and Eq. \ref{eq:eta-bulk} is reduced to:
\begin{equation}
  \label{eq:eta-bulk-2}
  \eta_{\mathrm{Bulk}} = \frac{1}{(1 - \hat{\eta}_{\mathrm{2D}} \gamma) (1 + \gamma)}
\end{equation}
The minimal value for $\eta_{\mathrm{Bulk}}$ when $\gamma>0$ is obtained by solving:
\begin{equation}
  \label{eq:eta-bulk-min}
  \frac{\partial \eta_{\mathrm{Bulk}}}{\partial \gamma}
  = \frac{2 \hat{\eta}_{\mathrm{2D}} \gamma + \hat{\eta}_{\mathrm{2D}} - 1}
  {(\gamma+1)^{2}(1 - \hat{\eta}_{\mathrm{2D}} \gamma)^{2}} = 0
\end{equation}
which gives that
$\eta_{\mathrm{Bulk}} \geq \frac{4
  \hat{\eta}_{\mathrm{2D}}}{(\hat{\eta}_{\mathrm{2D}} + 1)^{2}}$,
where the minimal value is taken at
$\gamma = \frac{1}{2}(\frac{1}{\hat{\eta}_{\mathrm{2D}}} - 1)$. Since
$\frac{4 \hat{\eta}_{\mathrm{2D}}}{(\hat{\eta}_{\mathrm{2D}} +
  1)^{2}}$ monotonically increases when
$0 < \hat{\eta}_{\mathrm{2D}} < 1$, we get the comparison between the
dielectric anisotropy indices between 2D materials and their bulk
counterparts:
\begin{equation}
  \label{eq:aniso-final}
\eta_{\mathrm{Bulk}} \geq \frac{4
  \eta_{\mathrm{2D}}}{(\eta_{\mathrm{2D}} + 1)^{2}} \geq
\eta_{\mathrm{2D}} 
\end{equation}
Since $\gamma$ is actually $2r_{0}^{\parallel}/L_{\mathrm{Bulk}}$, the
ratio between the in-plane screening length $r_{0}^{\parallel}$ and
inter-plane distance $L_{\mathrm{Bulk}}$, and in general
$r_{0}^{\parallel} \gg L_{\mathrm{Bulk}}$, we can conclude that the
case when $\eta_{\mathrm{Bulk}} = \eta_{\mathrm{2D}}$ only happens
when $\eta_{\mathrm{2D}}$ is much smaller than 1. Therefore the
separation between the 2D and 3D regimes in the Figure
\ref{main-fig:aniso} is explained.

\section{}

\section{Raw data from first principles calculations}
\subsection{Quantities from first principles calculation}
\label{sec:raw}

\input{raw_data}

\subsection{Density of States (DOS) plots for the materials studied}
\label{sec:DOS}

\input{DOS_figs}

\subsection{Bandstructures of 2D materials calculated}
\label{sec:bs}

\input{BS_figures}

\clearpage{}
\section*{}
\label{sec:ref}
\bibliography{ref}

\end{document}